---
title: "package_ideas"
author: "Lorenz Walthert"
date: "25 August 2016"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
source("projects/cfactor/cfactor.R")
source("projects/cfactor/cut.integer.R")
source("projects/cfactor/median.ordered.R")
source("projects/cfactor/quantile.ordered.R")

```

## Entire packages
## factorcontrol
A package that fixes some of the issues with factors, e.g. the ones mentioned [here](http://win-vector.com/blog/2014/09/factors-are-not-first-class-citizens-in-r/). 

### unmatched factors
A first function cfactor.R has been defined to give warnings when empty factors are created from strings or when existing strings are "not"translated" preserved. Essentially, it is a wrapper for `factor()`, c stands for enhanced control.
```{r}
string <- c("a", "b", "c")
factor(string, levels = c("b", "c", "d")) # the uggly outcome

cfactor(string, levels = c("b", "c", "d"))
```

### detect levels
`cfactor` should detect sequences in strings, and if existent, order the levels
```{r}
money <- c("EUR 1 - EUR 10", "EUR 11 - EUR 20", "Eur 21 - EUR 22") # accidentially already orderd, so current factor is fine
factor(money, ordered = T) 

money <- c("EUR 21 - EUR 22", "EUR 1 - EUR 10", "EUR 11 - EUR 20")  # the general case
```

```{r, eval = F}
cfactor(money) # cfactor detects levels using regular expressions
```

```{r, echo = F}
cat("[1] EUR 1 - EUR 10  EUR 11 - EUR 20 Eur 21 - EUR 22 \n Levels: EUR 1 - EUR 10 < EUR 11 - EUR 20 < Eur 21 - EUR 22
")
```

```{r, eval = F}
# To conclude, the following should be equal:
levels(cfactor(sample(money))) == levels(cfactor(money))
```
Other example strings
```{r, echo = F, include = T}
c(" 1 to 4", " 5 to 6") # properly separated
c(" 1 to 4", " 4 to 6") # not properly separated
c("from 1,000 to 2,000", "from 2000 to 4,000") # comma separated and 'from' and 'to'
c("4.0 / 4.1", "4.2 / 4.3") # point and slash separator
c("one minute", "three minutes", "1 hour")
```


### boundary labels for integer bins
using `cut` gives boundaries like [50, 55) [55, 60), even for integers. This is undesired in the case of integer values, so it should give boundaries like "50 - 55", "56 - 60". Possibly with the spacing being optional.

```{r}
random <- sample(100)
cut.default(random, breaks = seq(0, 100, by = 10))[1:10] # the uggly outcome with cut.default

class(random) # ok, we define an S3 method for integer

cut.integer(random, breaks = seq(0, 100, by = 10))[1:10]


```

### cut method for ordered factors
This method should allow to summarize levels of factors into fewer categories given some break points
```{r}
timeX <- c("very low", "rather low", "low", "medium", "upper medium", "high")
time <- factor(timeX, levels = unique(timeX), ordered = T)

cut.ordered(time, breaks = c("low", "medium", "high"))


```


### Median and quantile function for class ordered
```{r}
median(factor(money, ordered = T))
```

## Individual functions
### is.unique
A function that uses table() and returns `r T` if the vector contains unique values.

### %notin%
Opposite of %in%. Also seee adv-r, which, unwhich ect.

### table with exclude == NULL

### compare
A function that compares two elements and gives a summary

## from the net
use Stackoverflow to find the most popular questions with non-trivial solutions and develop functions to tackle these problems.
